### 2023-07-14 manuscript.tex
Explain details of voxel size and stacked scans [ci skip]
This build is based on
https://github.com/habi/EAWAG-manuscript/commit/a690fe23582f53c283f038e12d0f163b7ffa128a.

This commit was created by the following CI build and job:
https://github.com/habi/EAWAG-manuscript/commit/a690fe23582f53c283f038e12d0f163b7ffa128a/checks
https://github.com/habi/EAWAG-manuscript/actions/runs/5554735504

\emph{A preprint of this manuscript is available at bioRχiv with the \href{https://doi.org/10.1101/2023.03.30.534917}{doi:10.1101/2023.03.30.534917}.}
\emph{We submitted \href{https://github.com/habi/EAWAG-manuscript/releases/tag/v1.0}{\texttt{v1.0}} (\href{https://habi.github.io/EAWAG-manuscript/v/e1e2ef76a476174a4115937d77457037ddec95df/}{online version}) of the manuscript to \href{https://journals.plos.org/plosone/}{PLOS ONE}.}
\emph{The version you see here \href{https://github.com/habi/EAWAG-manuscript/compare/v1.0...a15f6a93f7f5b262f005c909a8fdcbdf8220da53}{has been updated} since submission.}

This manuscript
(\href{https://habi.github.io/EAWAG-manuscript/v/a15f6a93f7f5b262f005c909a8fdcbdf8220da53/}{permalink})
was automatically generated
from \href{https://github.com/habi/EAWAG-manuscript/tree/a15f6a93f7f5b262f005c909a8fdcbdf8220da53}{habi/EAWAG-manuscript@a15f6a9}
on July 13, 2023.

@@

Since the fish greatly varied in their length (specimen length varied between 6 cm and 18 cm), the voxel sizes of each of the acquired datasets also varies greatly.
We acquired datasets with (isometric) voxel sizes ranging from 3.5 to 50 μm.

Depending on the size of the specimen we set the X-ray source voltage to 50--80 kV and---depending on the voltage---to a current between 107 and 200 μA.
The X-ray spectrum was filtered either by an aluminum filter of varying thickness (either 0.25, 0.5 or 1 mm for increasing specimen size) before digitization to projection images or recorded in an unfiltered way (for smaller specimen).
Projection images spanning either 180° or 360° of sample rotation were acquired in angular steps of 0.1°, 0.15° or 0.2°, also depending on size of the fish.
In total we recorded 9.8 TB of projection images (\texttt{TIFF} and \texttt{*.iif} files, where the \texttt{*.iif} files are for the so-called alignment scans).

All the recorded projection images were subsequently reconstructed into virtual 3D stacks of axial \texttt{PNG} images spanning the regions of interest of each fish.
All the specimens were scanned with their mouths facing downward in the sample holder and rotating along their long axis.
We thus manually aligned each of the reconstructed datasets so that the lateral axis of the fish was horizontal in relation to the x and y direction of the reconstructed slices.
We reconstructed the projection images with NRecon (Version 1.7.4.6, Bruker microCT, Kontich, Belgium) with varying ring artifact and beam hardening correction values, depending on each fish (again, all relevant values are listed in the \protect\hyperlink{supplementary-materials}{Supplementary Materials}).
In total, this resulted in 1.5 TB of reconstruction images (a bit more than one million \texttt{*rec*.png} files).
On average, each of the scans we performed is made up of about 2700 reconstruction images.

@@

We acquired high resolution tomographic datasets of a large collection of cichlids (N=133), several tomographic scans were performed for each specimen.
The 372 acquired datasets were imaged over a wide-spanning range of voxel size (3.5--50 μm) permitting both the analysis of finest details we wanted to resolve (i.e.~containing only the oral and pharyngeal jaws) and having datasets (N=104) containing the whole head for principle components analysis and extraction of the otoliths.

>>>>

\emph{A preprint of this manuscript is available at bioRχiv with the \href{https://doi.org/10.1101/2023.03.30.534917}{doi:10.1101/2023.03.30.534917}.}
\emph{We submitted \href{https://github.com/habi/EAWAG-manuscript/releases/tag/v1.0}{\texttt{v1.0}} (\href{https://habi.github.io/EAWAG-manuscript/v/e1e2ef76a476174a4115937d77457037ddec95df/}{online version}) of the manuscript to \href{https://journals.plos.org/plosone/}{PLOS ONE}.}
\emph{The version you see here \href{https://github.com/habi/EAWAG-manuscript/compare/v1.0...a690fe23582f53c283f038e12d0f163b7ffa128a}{has been updated} since submission.}

This manuscript
(\href{https://habi.github.io/EAWAG-manuscript/v/a690fe23582f53c283f038e12d0f163b7ffa128a/}{permalink})
was automatically generated
from \href{https://github.com/habi/EAWAG-manuscript/tree/a690fe23582f53c283f038e12d0f163b7ffa128a}{habi/EAWAG-manuscript@a690fe2}
on July 14, 2023.

@@

Since the fish greatly varied in their length (specimen length varied between 6 cm and 18 cm), the voxel sizes of each of the acquired datasets also varies greatly.
We acquired datasets with (isometric) voxel sizes ranging from 3.5 to 50 μm.
Due to the geometric magnification employed by the Bruker SkyScan microCT machines, the voxel size of the resulting dataset is depending on the position of the specimen in the machine.
Natually, the field of view of the resulting dataset also depends on the chosen voxel size.
The chosen voxel size is thus scan-specific.
It was always chosen in a way that the region of interest to be images fit into the available lateral field of view of the microCT machine at the smallest necessary voxel size to resolve the structures of interest.
Since the region of interest (either oral or pharyngeal jaws or the complete skulls of the fish) often did not fit into a single field of view \emph{along} the anteroposterior axis of the fish, we often performed several tomographic scans along the rotation axis of the specimen in the machine (since the specimen were all scanned rotating around the anteroposterior axis).
The projection images from those so-called stacked scans are subsequently automatically merged to one stack of reconstructions images by the reconstruction software.

Depending on the size of the specimen we set the X-ray source voltage to 50--80 kV and---depending on the voltage---to a current between 107 and 200 μA.
The different voltages and currents were chosen according to the Bruker guidelines for ideal image acquisition.
The X-ray spectrum was filtered either by an aluminum filter of varying thickness (either 0.25, 0.5 or 1 mm for increasing specimen size) before digitization to projection images or recorded in an unfiltered way (for smaller specimen).
Projection images spanning either 180° or 360° of sample rotation were acquired in angular steps of 0.1°, 0.15° or 0.2°, also depending on size of the fish.
In total we recorded 9.8 TB of projection images (\texttt{TIFF} and \texttt{*.iif} files, where the \texttt{*.iif} files are for the so-called alignment scans).

All the recorded projection images were subsequently reconstructed into three-dimensional stacks of axial \texttt{PNG} images spanning the desired regions of interest of each specimen.
All the specimens were scanned with their mouths facing downward in the sample holder and rotating along their long axis.
We thus manually aligned each of the reconstructed datasets so that the lateral axis of the fish was horizontal in relation to the x and y direction of the reconstructed slices.
We reconstructed the projection images with NRecon (Version 1.7.4.6, Bruker microCT, Kontich, Belgium) with varying ring artifact and beam hardening correction values, depending on each fish (again, all relevant values are listed in the \protect\hyperlink{supplementary-materials}{Supplementary Materials}).
In total, this resulted in 1.5 TB of reconstruction images (a bit more than one million \texttt{*rec*.png} files).
On average, each of the scans we performed is made up of about 2700 reconstruction images.

@@

We acquired high resolution tomographic datasets of a large collection of cichlids (N=133), several tomographic scans were performed for each specimen.
The 372 acquired datasets were imaged over a wide-spanning range of voxel size (3.5--50 μm) permitting both the analysis of finest details we wanted to resolve (i.e.~containing only the oral and pharyngeal jaws) and having datasets (N=104) containing the whole head for principle components analysis and extraction of the otoliths.
In our microCT imaging setup the resulting voxel size is directly correlated with the sample diameter, i.e.~smaller specimen were scanned with the smallest voxel size, in the range of several μm.
The skulls of larger specimen were scanned with the largest voxel sizes, in the range of several tens of μm.
The chosen voxel size is thus specimen- or scan-specific and was always selected such that the desired region of interest fits into the available lateral field of view of the microCT machine at the minimally necessary voxel size to resolve the structures of interest.
